\reportsection{01}{Overview}{Eighteen Months of Patches, Read Back in Full}

\reppar{Since its first commit on December 26, 2024, \code{claude-desktop-debian} has done one thing at its core: take the \code{app.asar} out of Anthropic's Windows installer and make it behave like a native Linux application. The instrument for that is the \textbf{patch suite}: nine shell modules under \code{scripts/patches/} totaling 2,167 lines, plus 3,870 lines of injected JavaScript (\code{frame-fix-wrapper.js} at 997 lines, \code{cowork-vm-service.js} at 2,766, \code{claude-native-stub.js} at 107). Together they form thirteen patch units that rewrite minified upstream code by regex, stand in for Windows-only native binaries, and impersonate services the Windows app expects to find. The suite has been touched by \textbf{272 commits}, tracked \textbf{79 upstream version bumps}, and shipped in 147 tagged releases. This report widens the frame past the patch suite proper to the whole adaptation layer that surrounds it: the acquisition pipeline that fetches the upstream bytes, the launcher and doctor that decide runtime policy, and the packaging-time staging that shapes the shipped artifact, \textbf{eighteen units in all}.}

\reppar{On June 30, 2026, Anthropic shipped the first official Claude Desktop for Linux \code{.deb}, torn down in report \href{https://github.com/aaddrick/claude-desktop-debian/blob/main/docs/reports/CDL-ANT-0008_official-linux-teardown/claude-desktop-linux-teardown.pdf}{CDL-ANT-0008}. Two days later the project locked the decision to rebase v3.0.0 onto that official package, and the patch-necessity audit returned its verdict on the \code{app.asar} suite: \textbf{11 delete, 2 survivor candidates, 1 behavioral check, 1 parked} across the matrix's fifteen rows. A follow-on accessibility reassessment, re-verified against pristine official 1.18286.0, then annotated three of those calls with residual runtime risk the \code{.deb} bytes cannot settle, moving \code{frame-fix}, \code{wco-shim}, and \code{quick-window} to a \emph{verify} disposition without reversing any shipped code (section 21). Read under that reassessment and extended to the full adaptation layer this report covers, the settled disposition lands at the unit level as \textbf{eight units deleted outright, two survivors, four behavioral holds, one parked subsystem, and three reworked} rather than removed (the chassis, the launcher and doctor policy, and the acquisition pipeline). The deletions are not an admission of wasted work; in most cases the official build independently converged on the same fix the community had already derived, which makes the suite's history a record of correct diagnoses.}

\reppar{This report is the suite's dossier, written at the moment of its retirement. For every patch unit it answers four questions: what the patch does mechanically, why it was created (with the issues and pull requests that motivated it), how it was revised across eighteen months of upstream re-minification, and what the official build and the new \code{rebase/official-deb} build script do about it. Section 03 covers the shared chassis every patch runs on; sections 04\textendash 13 walk the ten \code{app.asar} units; sections 14\textendash 18 cover the surrounding layer this revision brings fully into scope, the acquisition pipeline, the launcher and doctor, and three packaging-time pieces; section 19 compresses everything into a fate matrix; section 20 closes with what the history means.}
