\reportsection{02}{Method}{Eighteen Dossiers, Each Through a Contrarian Gate}

\reppar{The findings here were produced by a multi-agent research workflow with adversarial verification at every step. One historian agent per patch unit reconstructed mechanism, origin, and revision history from the \code{main} branch: code was read via \code{git show main:\ldots}, origins were traced with pickaxe searches (\code{git log -S}) so that logic which migrated across the \code{build.sh} refactors is attributed to its true introduction rather than a file move, and every issue and pull request reference was resolved live against GitHub with \code{gh}. Each dossier then had to pass a \href{https://github.com/aaddrick/claude-desktop-debian/blob/main/.claude/agents/contrarian.md}{\textbf{contrarian gate}}: an adversarial reviewer that re-verified every cited issue number and title against reality, every verdict against the patch-necessity matrix in \code{docs/learnings/official-deb-rebase-verification.md}, every mechanism claim against the main-branch bytes, and every causal story against the commit or issue that states the cause. Dossiers that failed were revised and re-gated; all eighteen passed, the stubbornest after two rounds. A completeness critic swept \code{main} for machinery no dossier covered; its findings, three packaging-time pieces, together with two subsystems an earlier pass had scoped out (the acquisition pipeline, and the launcher and doctor policy), are the five units in sections 14\textendash 18. Each unit's \textbf{Affects} line reverse-looks-up its own issues and pull requests against a repo-wide environment catalog compiled from every issue and PR body, so distro, desktop, and compositor sensitivity is catalog-backed rather than asserted.}

\reppar{Two facts were additionally byte-verified for this report against the pinned official package (\code{claude-desktop\_1.17377.2\_amd64.deb}, SHA-256 \code{ec08d41\ldots212ab}, re-downloaded and hash-checked on July 3): the official tree ships its locale JSONs loose at \code{resources/*.json}, settling the fate of the i18n copy machinery, and it contains \textbf{no \code{claude-ssh} artifact anywhere}, which bears on the staged SSH helpers in section 14. Claims about the rebase side rest on reads of the checked-out \code{rebase/official-deb} working tree at \code{cafb4cc}.}

\methodbanner{\textbf{Scope.} This report covers the whole Linux adaptation layer on \code{main}: the acquisition pipeline that fetches the upstream bytes (section 14), the \code{app.asar} patch suite and its injected files that mutate or stand in for application code (sections 03\textendash 13), the launcher and doctor that set runtime policy (section 15), and the packaging-time staging that shapes the shipped artifact (sections 16\textendash 18). Where a verdict rests on evidence this project has not yet reproduced at runtime (the Quick Entry focus bug, the \code{mcpServers} overwrite, the in-app topbar question, the \code{claude-ssh} capability), the text says so plainly.}
