\reportsection{05}{Native binding}{The Native-Binding Stub: Eighteen Months of Pretending}

\unithead{Mechanism}

\reppar{Unlike the sed-based patches, \code{scripts/claude-native-stub.js} never touches the minified bundle. It is a \textbf{module-resolution shadow}: the bundle's \code{require("@ant/claude-native")} is satisfied by a drop-in \code{index.js} that the build copies into both load contexts, inside the asar (\code{scripts/patches/app-asar.sh}) and into the unpacked tree (\code{scripts/staging/electron.sh}), because the Windows \code{.node} binary at that path cannot load on Linux. No anchors, no identifier extraction; idempotency is \code{mkdir -p} plus an overwriting \code{cp}. The coupling: upstream null-guards only the module, never individual methods, so the stub must export everything upstream calls or a call throws during top-level execution.}

\unithead{Origin}

\reppar{The stub is as old as the repository, born in the initial commit of December 2024 as two heredocs in \code{build-deb.sh}: a frozen 19-entry \code{KeyboardKey} enum, a fixed \code{getWindowsVersion: () => "10.0.0"} spoof, and pure no-ops otherwise. It predates the tracker; the README credits k3d3's flake for the bindings insight (the specific key codes matching k3d3's work is inference, not a verified byte-trace). Growth came in waves: the \code{AuthRequest} class arrived to force login into the system browser (\#121, Google login spinner), PR \#143 (jacobfrantz1) retargeted the stub to the \code{@ant/} namespace, and PR \#228 (@milog1994) later made three methods genuinely functional.}

\begin{chronology}
2024-12 & \code{d8f4bbe} & Origin: initial commit embeds the stub twice as heredocs in \code{build-deb.sh}, targeting the un-namespaced \code{node\_modules/claude-native/index.js}; \code{KeyboardKey} enum, \code{getWindowsVersion} spoof, pure no-ops, no \code{getWindow()}, no \code{AuthRequest}. \\
2025-11 & \#121, \code{ae6d3a3} & The \code{AuthRequest} stub (\code{isAvailable()} false, throwing \code{start()}) forces Google login to fall back to the system browser, fixing the indefinite loading spinner; direct commit to main, no associated PR. \\
2025-11 & PR \#143, \code{5c5eb39} (@jacobfrantz1) & Both heredoc destinations retargeted from \code{node\_modules/claude-native/} to \code{node\_modules/@ant/claude-native/}, tracking upstream 1.0.1307's package rename; same-day follow-up \code{1405e1c} fixed a missing \code{mkdir -p} for the new namespaced directory. \\
2026-01 & PR \#180/\#181, \code{4bf5986} & Heredoc extracted to the standalone \code{scripts/claude-native-stub.js}, duplicated second definition removed; one source file thereafter, copied to both destinations. \\
2026-02 & PR \#228, \code{83cbb9a} (@milog1994) & \code{getIsMaximized}, \code{flashFrame}, and \code{setProgressBar} made genuinely functional via a new \code{getWindow()} helper, fixing \#149 (KDE Plasma attention flash). \\
2026-02 & PR \#232, \code{2245808}, \code{75841f0} & Hardening series: destroyed/invisible windows filtered out of the \code{getWindow()} fallback, then the \code{isVisible()} filter deliberately dropped so \code{flashFrame()} keeps working on minimized windows, its primary use case. \\
2026-04 & \code{ff4821e} & Pure move: injection logic lands in \code{scripts/patches/app-asar.sh} (\code{patch\_app\_asar}) and \code{scripts/staging/electron.sh} (\code{finalize\_app\_asar}); no behavior change. \\
2026-06 & \#729, PR \#737, \code{295d71b} & Upstream \(\geq\) 1.13576.0 calls \code{readRegistryValues()} and \code{getWindowsElevationType()} unconditionally at startup; the old stub threw, an empty \code{uncaughtException} handler swallowed it, and the app hung windowless. PR \#737 consolidated \#734 (@chrisw1005) and \#730 (@colonelpanic8) into five neutral registry no-ops plus \code{tests/claude-native-stub.bats}, crediting both authors. \\
\end{chronology}

\methodbanner{\textbf{Fate under the rebase.} \bdel{} The matrix verdict is unconditional: ``\textbf{delete} \ldots Real Rust NAPI ELF at \code{resources/app.asar.unpacked/node\_modules/@ant/claude-native/claude-native-binding.node}.'' That binding is a 1.65 MB NAPI-RS Rust ELF with genuine X11 input injection (\code{enigo} plus \code{x11rb}/XTEST), \code{rustix}/\code{openat2} containment, and \code{SO\_PEERCRED} peer auth, capabilities a JS stub could never fake; \code{tools/patch-necessity-audit.sh} confirms the ELF and reports \code{not-needed}. At \code{d9cef9e} the stub and its bats suite are deleted, neither injection site survives, and the wholesale \code{ar p | tar} extraction plus a repack that hard-fails on unpacked-set divergence ships the binding untouched. Known-stale by design: \code{tests/test-artifact-common.sh} still asserts \code{index.js} exists; that rework belongs to @sabiut, with the \code{test-artifacts} CI jobs expected red until it lands.\\[3pt]\textbf{Affects.} All Linux; a module-resolution shadow, DE-agnostic.}
