\reportsection{07}{Tray}{Tray: Five Anchor-Rot Cycles Toward Upstream\textquotesingle s Own Design}

\unithead{Mechanism}

\reppar{\code{scripts/patches/tray.sh} is four functions against the minified main-process bundle, three of them tray proper. \code{patch\_tray\_menu\_handler} finds the rebuild function through the \code{on("menuBarEnabled",()=>\{FUNC()\})} anchor, rewrites it \code{async}, injects a \textbf{trailing-edge mutex} (\code{\_running}/\code{\_pending}, 1500ms) so a rebuild request arriving mid-flight is remembered and the final \code{nativeTheme} value wins, and adds a 250ms post-\code{destroy()} sleep so the StatusNotifierItem can unregister before a new \code{Tray} registers. \code{patch\_tray\_icon\_selection} rewrites the hardcoded \code{TrayIconTemplate.png} assignment into a \code{shouldUseDarkColors} conditional (the icons themselves were made opaque at build time, since the originals are macOS \textasciitilde{}20\%-opacity templates Linux never colorizes). \code{patch\_tray\_inplace\_update} extracts five minified locals and splices a fast-path ahead of the destroy-and-recreate block: if a tray exists and the feature is enabled, \code{setImage} plus \code{setContextMenu} in place, \code{return}. The fourth resident, \code{patch\_menu\_bar\_default}, captures the preference variable off the \code{"menuBarEnabled"} string literal and rewrites the gate's \code{,!!VAR)} coercion to \code{VAR!==false}, so a missing config key means tray on rather than \code{!!undefined} meaning tray off. Two \code{\_common.sh} helpers, \code{extract\_electron\_variable} and \code{fix\_native\_theme\_references}, serve the family; the latter exists to repair an \emph{upstream} minifier bug class where the bundle referenced \code{oe.nativeTheme} while electron was bound to \code{Ae}.}

\begin{codeblock}
if(TRAY && ENABLED!==false){
  TRAY.setImage(EL.nativeImage.createFromPath(PATH));
  process.platform!=="darwin" && TRAY.setContextMenu(BUILDER());
  return
}
\end{codeblock}

\unithead{Origin}

\reppar{The tray arc opened in November 2025 with \#135 (Wayland DBus "already exported", a non-functional tray menu): commit \code{1bb05df} added the mutex and delay with hardcoded names, and \code{c660bf1} made them dynamic the same day. The menubar-default patch is the family's quieter passenger, born in \code{7bd2f38} against \#218's twin complaint: the \code{!!undefined} gate plus updaters dropping the key from \code{config.json}. Across the life of the file the suite absorbed \textbf{five anchor-rot re-derivations}, called out below.}

\begin{chronology}
2025-11 & \#135, \#138 \code{1bb05df}; \#139 \code{c660bf1} & Origin: \code{async} rewrite, mutex, and post-destroy delay against the DBus double-export; hardcoded minified names made dynamic the same day. \\
2026-01 & \#158, \code{bed0cc3} & Anchor rot 1: upstream 1.0.3218 added an early return before the first \code{const}; the mutex insertion re-anchored on the function brace (@lizthegrey). \\
2026-01 & \#163, \code{e5d58f2}, \code{6916a9e} & Icon selection plus build-time opacity processing against invisible template icons; the light/dark logic inverted on day one and corrected the next; mutex widened to 1500ms and the post-destroy delay to 250ms after a first attempt (\#164) leaned on a Chromium flag that "doesn't exist in Electron/Chromium". \\
2026-02 & \#218, \#219, \code{7bd2f38} & Upstream's own \code{oe}/\code{Ae} minifier bug in v1.1.2321; birthed both \code{fix\_native\_theme\_references} and the \code{menuBarEnabled} default patch. \\
2026-02 & \#252, \code{546f845} & Anchor rot 2: v1.1.4088 minified the electron binding to \code{\$e} and \code{\textbackslash w}-only capture spliced code mid-name, a launch-time SyntaxError; anchors widened to accept dollar identifiers. \\
2026-04 & \#515, \code{cf2b0fc} & The in-place \code{setImage}/\code{setContextMenu} fast-path ahead of destroy-and-recreate: the destroy-plus-delay race is structural on KDE Plasma, the old SNI staying registered while the new one appears, two icons until logout (@IliyaBrook; diagnosis preserved in \code{docs/learnings/tray-rebuild-race.md}). \\
2026-05 & \#625, \#627 \code{6219d5a}; \#644 \code{b40441c} & Anchor rot 3: 1.8089.1's \code{i\$A} handler beat PCRE \code{\textbackslash w}; \code{[\$\textbackslash w]+} then became convention repo-wide in the \#644 regex audit (@typedrat). \\
2026-05 & \#656, \code{e38066e} & Anchor rot 4: 1.9255.0 reshuffled declarations; the tray variable re-anchored on the \code{Tray} constructor literal. \\
2026-06 & \#679, \#680, \code{e13e331} & Mutex converted to trailing-edge after dark-GNOME startups latched a stale \code{shouldUseDarkColors=false} and stuck the icon black (@LiukScot). \\
2026-06 & \#750, \code{6091615}, \code{eb12ad8} & Anchor rot 5: the 1.13576+ "yukonSilver" rebuild refactor silently re-armed the \#515 race through green CI; the resolver learned to skip the \code{setContextMenu(null)} decoy and assert exactly one \code{TRAY.destroy()} anchor, and the menubar patch learned the new upstream defaults map, warning loudly on a genuine anchor miss instead of skipping. \\
\end{chronology}

\methodbanner{\textbf{Fate under the rebase.} \bdel{} Both rows are unconditional deletes against official 1.17377.2. The tray verdict: "Official rebuild takes an in-place \code{setImage} branch keyed on icon-path change; \code{Tray.destroy()} only runs when the user disables the tray. No SNI re-registration gap exists," and the Linux branch "natively selects \code{TrayIconLinux(-Dark).png}" (the \code{TrayIconTemplate.png} anchor survives only in the macOS branch). Upstream independently converged on exactly what \code{patch\_tray\_inplace\_update} injected, purpose-made opaque icons included: a validation of the community diagnosis. The menubar row: "Defaults map ships \code{menuBarEnabled:!0}," the same \code{Bpt=\{menuBarEnabled:!0\}} bytes first captured in \#750, so an unset preference already resolves true. Commit \code{d9cef9e} on \code{rebase/official-deb} deleted \code{tray.sh}, \code{\_common.sh}, \code{tests/tray-patches.bats}, and the ImageMagick icon pipeline; \code{active\_patches} carries no tray entry, and \#750 is mooted by deleting the patch it tracks. One hedge from the verification record: \code{fix\_native\_theme\_references} has no explicit matrix row or audit probe, so its deletion rests on tray.sh's wholesale removal rather than a byte-level check that 1.17377.2 is free of the \#218-class bug.\\[3pt]\textbf{Affects.} KDE Plasma and GNOME (SNI-tray desktops); Wayland and X11, where the DBus/SNI re-registration race lived.}
