\reportsection{12}{Topbar}{The WCO Shim: One Commit, Zero Revisions}

\unithead{Mechanism}

\reppar{\code{patch\_wco\_shim} (\code{scripts/patches/wco-shim.sh}) is the only patch in the suite that rewrites nothing: it performs a pure prepend, inlining \code{scripts/wco-shim.js} at the top of \code{app.asar.contents/.vite/build/mainView.js}, the BrowserView preload, guarded by a \code{grep -q '\_\_claude\_wco\_shim'} idempotency check and a hard \code{exit 1} if the target file is missing. The inline (rather than a \code{require}) is deliberate: sandboxed preloads can only require a fixed allowlist of modules, and a relative require aborts the entire preload, taking \code{desktopBootFeatures} down with it. At runtime the shim, Linux-only and disabled when \code{CLAUDE\_TITLEBAR\_STYLE} is \code{native}, injects a main-world script via \code{webFrame.executeJavaScript}. Its components: a diagnostic native-state probe, defensive \code{navigator.windowControlsOverlay} and \code{matchMedia} shims, a \code{draggable}-stripping className intercept, an event nudge, and \textbf{the one load-bearing piece}, a page-side \code{navigator.userAgent} getter that appends \code{" Windows"} so the remote claude.ai bundle's \code{isWindows()} gate flips and React renders the desktop topbar (\code{data-testid="topbar-windows-menu"}). The HTTP request UA is unchanged, "so analytics and anti-bot fingerprints stay honest."}

\unithead{Origin}

\reppar{The unit was born whole in \code{5c8191e} (May 2026, \#538, hybrid titlebar mode): the topbar was never in \code{app.asar}, since claude.ai's remote React bundle renders it behind four gates, of which only Gate 3, the \code{/(win32|win64|windows|wince)/i} UA regex, fails on Linux (documented in \code{docs/learnings/linux-topbar-shim.md}). \#127 (speleoalex's native-decorations PR, November 2025) had forced \code{frame: true}, "which definitively hid the topbar," and the upstream frameless-plus-WCO route was independently broken by a Chromium-level implicit drag region on frameless Linux windows (Bug C) that left topbar buttons unclickable. Hybrid mode, system frame plus page-side UA spoof, was the resolution; the same commit deleted the predecessor \code{patch\_titlebar\_detection}, which since April 2025 had stripped the \code{!} from \code{if(!isWindows \&\& isMainWindow)} in bundled renderer assets, the fix that had closed \#45 (missing hamburger menu, boyonglin), an attribution inferred from timing and the issue's closing comment. The zero-revision record that follows is the exception that proves the arms-race rule chronicled in the chassis section: because the shim overrides stable web-platform APIs instead of grepping minified identifiers, upstream re-minification could never break it.}

\reppar{The defensive layer also drove one contribution to Electron itself rather than a workaround in the shim: chasing why \code{matchMedia("(display-mode: window-controls-overlay)")} stayed false while every other WCO signal reported active, aaddrick filed \href{https://github.com/electron/electron/issues/51393}{electron\#51393} (Electron never overrides \code{GetDisplayMode()}, so Blink's media-query evaluator always sees \code{kBrowser}) and opened the fix in \href{https://github.com/electron/electron/pull/51396}{electron\#51396}. Both are open as of this writing, the pull request carrying backport labels for Electron 41 through 44.}

\begin{chronology}
2025-04 & \#45, \code{d6ed2c8} & Predecessor: \code{patch\_titlebar\_detection} stripped the \code{!} from \code{if(!isWindows \&\& isMainWindow)} in bundled renderer assets, closing \#45 (missing hamburger menu, boyonglin; attribution inferred from timing plus the issue's closing comment). \\
2026-05 & \#538, \code{5c8191e} & Origin: hybrid titlebar mode; the shim's introduction and \code{patch\_titlebar\_detection}'s deletion landed in the same commit. Pickaxe (\code{git log -S}) confirms both \code{wco-shim.sh} and \code{wco-shim.js} carry a single-commit history on main: no anchor rot, ever. \\
2026-05 & \#538 (in-thread) & Unreproduced regression: lukedev45 reported partial topbar render on OmarchyOS plus Hyprland; aaddrick failed to reproduce across four scenarios, @typedrat confirmed working on NixOS Hyprland (also GNOME Wayland and Sway in-thread), and no follow-up issue was filed. \\
\end{chronology}

\methodbanner{\textbf{Fate under the rebase.} \bdel{} (local WCO) \bverify{} (remote UA gate) The accessibility reassessment splits the verdict. \code{probe\_wco\_shim()} returns \code{not-needed}, but the audit qualifies it ``(mainView refs: 0)'': the official \code{mainView.js} has no \code{windowControlsOverlay}/\code{isWindows} gating and the main window omits \code{frame} entirely, so Bug C cannot arise and the local frameless/WCO/matchMedia defensive half is genuinely dead (delete stands). That check reaches only the \emph{local} bundle. The one load-bearing piece, the page-side \code{navigator.userAgent} spoof that flips claude.ai's remote \code{isWindows()} UA regex so the desktop topbar (\code{data-testid="topbar-windows-menu"}) renders at all, depends on server-delivered JS that \code{.deb} bytes cannot reach, so that half moves to \bverify{}: open check \textbf{WCO-1} (topbar DOM presence on a rebased AppImage or RPM build, \emph{not} the official \code{.deb}) decides delete versus a UA-override-only survivor. This annotates residual risk; the code is already gone. Both files were deleted at \code{d9cef9e}, the titlebar machinery left the launcher at \code{cafb4cc}, and \code{doctor.sh} now warns that \code{CLAUDE\_TITLEBAR\_STYLE} ``is set but no longer honored since the v3.0.0 rebase.'' If the remote bundle still Windows-gates the topbar, v3.0.0 silently loses the hamburger/search/nav bar that v2.x hybrid mode delivered; per the tracking plan, ``if missing, it's an upstream report, NOT a shim revival.''\\[3pt]\textbf{Affects.} All Linux equally: it overrides stable web-platform APIs, so DE- and compositor-agnostic by construction.}
