\reportsection{13}{Config writes}{Config Writes: One Bug, Three Patches, a Split Verdict}

\unithead{Mechanism}

\reppar{\code{scripts/patches/config.sh} carries three functions, all operating on \code{app.asar.contents/.vite/build/index.js}. \code{patch\_config\_write\_merge} anchors on the developer log string that survives every minifier pass, the central write site \code{await WRITE\_FN(PATH, CONFIG), LOGGER.info("Config file written")}, extracts the minified names with three chained \code{grep -oP} passes, then uses \code{node -e} to prepend a merge statement: \textbf{re-read the config from disk on every write}, take on-disk \code{mcpServers} as the base, and let in-memory entries override matching keys, so externally added servers survive writes made from the app's stale in-memory cache. \code{patch\_asar\_trusted\_folder\_guard} anchors on the unminified \code{async addTrustedFolder(} declaration and injects \code{if(PARAM.endsWith(".asar"))return;} at the body head. \code{patch\_asar\_additional\_dirs\_guard} filters every \code{--add-dir} dispatch loop (for-of and \code{.forEach} variants) plus a best-effort session-restore self-heal keyed to the \code{"Filtering out deleted folder from session"} string.}

\begin{codeblock}
'await \K[$\w]+(?=\([$\w]+,\s*[$\w]+\)\s*,\s*[$\w]+\.info\("Config file written"\))'
\end{codeblock}

\unithead{Origin}

\reppar{The unit was born from \#400 ("claude\_desktop\_config.json being reset continuously", opened April 2026 by @davidcim): every start or mode switch rewrote the config with stale content, making MCP servers impossible to add; the same doctor paste showed \code{app.asar} recorded in \code{localAgentModeTrustedFolders}, the amplifying symptom the trusted-folder guard targets. Commit \code{364147e} (PR \#643) created both functions, diagnosing an upstream writer that \textbf{caches parsed config and never re-reads disk before writing}. Days later @beneshengineering filed \#649 (asar reaching \code{--add-dir} despite the \#640 cowork guard): corrupted pre-\#640 sessions survived restore via Electron's ASAR VFS shim and crashed local agent mode under bundled Claude Code 2.1.111, so PR \#650 added the additional-dirs guard at "the single convergence point for ALL code paths that feed additionalDirectories".}

\begin{chronology}
2026-05 & \#400, \code{364147e} (PR \#643) & Origin: created \code{patch\_config\_write\_merge} and \code{patch\_asar\_trusted\_folder\_guard} against the stale in-memory cache overwriting \code{mcpServers} on every config write (@aaddrick). \\
2026-05 & \#649, \code{4451694} (PR \#650) & Added \code{patch\_asar\_additional\_dirs\_guard}: filters every \code{--add-dir} dispatch loop and self-heals session restore, after corrupted pre-\#640 sessions kept reaching \code{additionalDirectories} despite the existing guards (@beneshengineering). \\
2026-06 & \#674/\#685, \code{2ede75d} & Anchor rot: upstream 1.10628.0 folded the trusted-folder guard's log-line anchor into a comma expression (the trailing \code{);} became \code{),}), hard-failing the build; @maplefater's PR \#685 re-anchored on the method declaration itself, superseding @mhentschke's identical \#674, closed unmerged two minutes after \#685 landed. \\
2026-06 & \#718/\#722/\#723, \code{5e4f26b} & Anchor rot: upstream 1.12603.1 bundled the Claude Code SDK per-panel, producing two identical \code{--add-dir} dispatch loops and tripping the single-match assertion (\#718, nix build failure, reported by @fdnt7); @typedrat's PR \#723 reframed the invariant as "every unfiltered --add-dir dispatch must filter .asar paths" via a global replace, with @marveon's competing \#722 credited for the diagnosis. \\
\end{chronology}

\methodbanner{\textbf{Fate under the rebase.} Split. Both \code{.asar} guards: \bdel{} "delete", per the matrix, "\code{addTrustedFolder(o)} present without a \code{.asar} guard, but same reasoning as above: no on-disk \code{.asar} argv path exists on Linux": the official launcher is a bare ELF symlink, the reasoning first stated on main in ab17b69. The mcpServers merge: \bverify{} "verify behaviorally", the suite's one open question. The \code{Config file written} anchor is byte-intact on official 1.17377.2 (audited 2026-07-02, reproducible via \code{tools/patch-necessity-audit.sh}), which proves structural continuity of the writer, not that the stale-cache bug persists. On \code{rebase/official-deb} (d9cef9e), \code{config.sh} shrank from 296 to 104 lines, guards and \code{tests/config-patches.bats} gone; the merge survives in the file but is kept \textbf{unwired}, absent from \code{app-asar.sh}'s \code{active\_patches}, pending a reproduction of \#400 against a live official install, with an upstream filing due either way. Deferred, not decided.\\[3pt]\textbf{Affects.} All Linux; the config-write merge and argv guards, DE-agnostic.}
