\reportsection{18}{Sandbox}{The Sandbox Shims: What the SUID Bit Survives On}

\unithead{Mechanism}

\reppar{The deb postinst (\code{scripts/packaging/deb.sh}) locates \code{chrome-sandbox} under the installed package tree and runs \code{chown root:root} plus \code{chmod 4755} on it, warning rather than failing if the file is missing. The comment states the reason directly: the official \code{data.tar.xz} records the SUID bit, but the project's non-root \code{ar | tar} extraction (\code{scripts/setup/official-deb.sh}, which deliberately avoids \code{dpkg-deb} so RPM-family and Arch hosts can build) strips it, so the postinst restores what extraction lost. \code{scripts/packaging/rpm.sh} takes a different shape: rather than a \code{\%post} scriptlet, \code{chmod 4755} is baked into the buildroot before \code{\%install}'s \code{\%files} walk, ordered to run after the blanket \code{find ... -exec chmod u=rwX} permission-normalization pass so the 4755 bit survives it. \code{scripts/packaging/appimage.sh} passes \code{--no-sandbox} unconditionally through \code{build\_electron\_args()}, because a FUSE-mounted, user-mounted AppImage cannot host a root-owned, 4755-mode helper the way an installed package can: there is no SUID path available inside the AppDir, so the flag is the only option rather than a workaround. On \code{main}, \code{nix/claude-desktop.nix} repoints the stock nixpkgs electron-wrapper's \code{CHROME\_DEVEL\_SANDBOX} at a co-located copy of \code{chrome-sandbox} (a path-correctness fix, not a permission-loss fix, since Nix builds never go through \code{ar | tar}); on the rebase branch that derivation is currently a stub.}

\begin{codeblock}
chown root:root "$SANDBOX_PATH"
chmod 4755 "$SANDBOX_PATH"
\end{codeblock}

\unithead{Origin}

\reppar{Pickaxe resolves each shim to a distinct commit, and none share an origin despite superficially similar mechanics. The deb reassert traces to \code{d7d4695} (2025-03-30, direct commit, no PR), reverting a same-day, ten-minutes-earlier \code{ed2ac2d} that had hardcoded \code{--no-sandbox} into the deb launcher to fix a Google-login crash, citing \code{electron/electron\#17972}. The AppImage flag traces to \code{1f13d7a} (2025-04-02, direct commit): unlike the deb flag, it was never superseded, since a portable AppImage has no postinst-style mechanism to fall back to. The Wayland-conditional \code{--no-sandbox} for deb traces to \code{3e43708} (\#97, @delorenj), bundled with GPU flags to fix a \code{Trace/breakpoint trap} crash attributed to dmabuf/EGLImage binding errors on GNOME and KDE Plasma Wayland. The RPM shim traces to \code{86a1822} (2026-01-24): multi-distro support ported the deb's already-abandoned first-draft \code{\%post} \code{chown}/\code{chmod} pattern into \code{rpm.sh}, not its later, current shape. That RPM shim then ran its own two-round saga: \#539 (@qtlin) reported that a \code{\%post} \code{chmod} silently no-ops under \code{--noscripts} or layered \code{rpm-ostree} images, shipping a non-SUID \code{chrome-sandbox}; PR \#595 (@JoshuaVlantis) moved the bit into \code{\%attr(4755,\ root,\ root)} in \code{\%files}, which introduced a new \code{rpmbuild} "File listed twice" warning caught as \#609 (aaddrick) and fixed by PR \#610 (@JoshuaVlantis), which baked the chmod into the \code{\%install} buildroot instead. Separately, the AppArmor \code{userns} arc opened with \#351 (@hfyeh, Ubuntu 24.04 blocking Cowork's \code{bwrap}), got a manual, system-wide doc workaround in PR \#434 (@RayCharlizard) that \#542 later flagged as over-broad, then PR \#687 (@diarized) automated a scoped profile for the Electron binary itself against a deb-on-X11-only \code{credentials.cc} crash, explicit that the sandbox "stays enabled, never \code{--no-sandbox} on the deb," and PR \#694 (same author, same day) added the structurally identical scoped profile for \code{/usr/bin/bwrap}, superseding \#434's manual fix.}

\begin{chronology}
2025-03 & \code{cc57c39}; \code{ed2ac2d}; \code{d7d4695} & README documents \code{--no-sandbox} as a manual step; deb launcher hardcodes it, citing electron/electron\#17972; reverted the same day by the postinst \code{chown root:root}/\code{chmod 4755} reassert (origin of the deb shim). \\
2025-04 & \code{1f13d7a}; \code{83d302a} & Origin of the AppImage shim: \code{--no-sandbox} added permanently for FUSE; deb packaging logic extracted into its own script. \\
2025-06 & \#97, \code{3e43708}; \code{bf6c5bf} & Origin of the Wayland-deb shim: \code{--no-sandbox} bundled with GPU flags against a \code{Trace/breakpoint trap} crash on GNOME and KDE Plasma Wayland (@delorenj); merged 2025-08-04. \\
2026-01 & \code{86a1822}; \code{ee5e090}, \code{424e17b} & Origin of the RPM shim, porting the deb's abandoned first-draft \code{\%post} pattern; per-branch \code{--no-sandbox} calls consolidated into \code{build\_electron\_args()}. \\
2026-03 & issue \#282, \code{99d91ac}; issue \#351 & Wayland branch extended from \code{deb} to \code{deb || nix}; @hfyeh reports Ubuntu 24.04's userns restriction blocking Cowork's \code{bwrap}. \\
2026-04 & PR \#368 (fixes \#316), \code{a326ea2}; PR \#434; issue \#539 & Origin of the Nix shim, a side effect of the isPackaged/resourcesPath fix; @RayCharlizard ships a manual, system-wide \code{bwrap} AppArmor profile; @qtlin reports the RPM \code{\%post} chmod is scriptlet-fragile. \\
2026-05 & \#539\,\arrow\,PR \#595, \code{15813ca}, \code{cf08571}, \code{c9df9e2} & RPM SUID moved from \code{\%post} chmod to \code{\%attr(4755,\ldots)} in \code{\%files} (@JoshuaVlantis); merged 2026-05-14. \\
2026-05 & issue \#609\,\arrow\,PR \#610, \code{ba8ffa1}, \code{a04ed9e} & RPM SUID moved again, from the explicit \code{\%attr} line to a buildroot \code{\%install} chmod, silencing "File listed twice" (@JoshuaVlantis); merged 2026-05-24. \\
2026-06 & PR \#695, \code{5c43ccd} & Blanket install-tree permission normalization for the Cowork daemon launch fix (@caidejager), sequenced to run before the chrome-sandbox-specific chmod so 4755 survives. \\
2026-06 & PR \#687, \code{12cc726}; PR \#694, \code{50dd1f0} & Scoped AppArmor userns profile automated for the Electron binary (deb-on-X11 crash fix); a structurally identical profile for \code{/usr/bin/bwrap} the same day, superseding \#434's manual fix (@diarized). \\
2026-07 & \code{d9cef9e} & Phases 1+2: paths updated to the official package root (no more \code{node\_modules/electron/dist}); AppArmor profile's attached binary path updated; Nix derivation replaced with a stub. \\
2026-07 & \code{cafb4cc} & Phases 4+5: doctor's chrome-sandbox and User-namespaces checks moved to the official layout; \code{tools/chromium-switch-smoke.sh} added to lock the \code{--no-sandbox} scenario matrix. \\
open & issue \#617 & @sabiut's \code{--noscripts} RPM regression-test gap remains unresolved. \\
\end{chronology}

\methodbanner{\textbf{Fate under the rebase.} \bsurv{} The shims survive because the official \code{data.tar.xz} records \code{chrome-sandbox} SUID (\code{-rwsr-xr-x root/root}), but the project's non-root \code{ar | tar} extraction strips that bit, so the postinst \code{chown root:root}/\code{chmod 4755} reassert ports over unchanged in logic as the mechanism that restores what extraction lost, its path merely updated from \code{node\_modules/electron/dist/chrome-sandbox} to the bare package-root \code{chrome-sandbox} in the official layout. The scoped Electron-binary AppArmor \code{userns} profile survives alongside it, re-attached to the bare official ELF, sitting next to the official postinst's own AppArmor and apt self-registration behavior in that same install step. \code{rpm.sh}'s buildroot chmod and \code{appimage.sh}'s FUSE \code{--no-sandbox} survive unchanged in effect; \code{doctor.sh}'s chrome-sandbox and User-namespaces checks were reworked to the new paths, and a new \code{tools/chromium-switch-smoke.sh} locks the \code{--no-sandbox} scenario matrix in CI. What does not survive: the structurally identical scoped AppArmor profile for \code{/usr/bin/bwrap} (\#694), parked because the official Cowork backend is \code{coworkd} plus QEMU/KVM rather than \code{bwrap}; \code{nix/claude-desktop.nix} is presently a stub, so the Nix \code{CHROME\_DEVEL\_SANDBOX} repoint's fate is open rather than settled.\\[3pt]\textbf{Affects.} RPM SUID payload and Ubuntu 24.04+ unprivileged-userns restriction (AppArmor); AppImage forces \code{--no-sandbox} under FUSE.}
