\reportsection{20}{Closing}{A Record of Correct Diagnoses}

\reppar{Read end to end, the suite's history is dominated by one pattern: \textbf{the community found the bug first, and upstream eventually agreed}. The tray unit spent eight months converging on an in-place \code{setImage} design that the official build then shipped natively; the frame fix forced the system titlebar that the official window simply has; the menu-bar default, the Linux \code{getHostPlatform} branches, the pty binary, and the disabled updater all reappear in the official package as first-party decisions. Deleting the redundant units under the rebase discards code, not knowledge: the diagnoses were validated by the strongest reviewer available, the vendor's own engineers arriving at the same answers.}

\reppar{What survives is exactly the set of problems upstream has not solved. \code{quick-window.sh} stays because the Electron-on-KDE stale-focus anchor is still present in official bytes; \code{org-plugins.sh} stays because the official platform switch still returns \code{null} on Linux, leaving MDM-managed plugin marketplaces dead upstream; the \code{mcpServers} merge waits unwired on a behavioral reproduction of \#400; and the Cowork daemon is parked, not deleted, because a bwrap fallback for non-KVM hosts remains genuinely useful and now means speaking \code{coworkd}'s \code{SO\_PEERCRED} protocol, a 3.1-sized investigation. The survivor suite also keeps the arms-race methodology alive: \code{docs/learnings/patching-minified-js.md} governs two patches now instead of thirteen, against an upstream that shipped 1.18286.0 one day after 1.17377.2.}

\reppar{The history also leaves the project a to-do list it can now file where it belongs. The suite's remaining gaps convert into \textbf{upstream reports against a first-party Linux target}: the missing org-plugins Linux case, the possible loss of the in-app topbar behind the remote bundle's Windows gate, the \code{StartupWMClass} mismatch, the hardcoded OVMF probe list, and the stdio MCP double-spawn already validated in CDL-ANT-0008. Open items on this report's own ledger: the KDE Plasma reproduction that decides \code{quick-window.sh}, the \#400 reproduction that decides the config merge, and a runtime answer to where, if anywhere, the official Linux build carries the \code{claude-ssh} capability. And the rebase clarifies the repository's purpose rather than ending it: Anthropic ships Linux as a single Debian-family \code{.deb}, so the mission now is the long tail, repackaging the official build as RPM, AppImage, and Nix derivation for the distributions the first-party channel does not reach. The patch suite ends as it lived: not as a fork of Claude Desktop, but as the best available record of what it actually took to run one on Linux.}
