\clearpage
\reportsection{21}{Addendum}{An Accessibility Reassessment, Re-Verified Against 1.18286.0}

\reppar{This section is an \textbf{addendum}, not a nineteenth unit: the eighteen-unit spine and every count in the overview and the fate matrix stand as written. After the retirement audit was fixed against the pinned official 1.17377.2 bytes, a second pass (\textbf{accessibility reassessment, July 3, 2026}) re-scored every verdict under a single lens, \textbf{make the Linux build work out-of-the-box on the widest range of the environments users actually report}, and ground-truthed the result against a freshly fetched pristine official \textbf{1.18286.0} \code{.deb} (sha256 \code{8f314ad1\ldots0536}, three releases newer than the pin). The shipped code did not change; what changed is how much residual risk three of its deletions and survivor calls are now understood to carry.}

\unithead{Three flips, one firm}

\reppar{The lens moved exactly three verdicts, all in the same direction, from a confident \bdel{}/\bsurv{} to \bverify{}, because in each case the load-bearing evidence lives where \code{.deb} bytes cannot reach. \code{frame-fix} splits: the frame-core, titlebar-mode, and \code{autoUpdater} no-op slices are byte-moot and stay deleted, but the roughly eighteen accreted Electron-runtime fixes track \emph{unfixed upstream} bugs (\#416 hover-raise, \#605 sleep inhibitor, \#128 \code{openAtLogin}, \#623 quit hatch), and \code{tools/patch-necessity-audit.sh} returns \code{check} on 1.18286.0 (``\code{frame:!1} occurs 3x''), not a clean delete. \code{wco-shim} splits the same way: the local \code{mainView.js} WCO half is dead (audit \code{not-needed}, ``mainView refs: 0''), but the load-bearing page-side \code{isWindows()} UA spoof depends on claude.ai's server-delivered bundle. \code{quick-window} moves from survivor candidate to \code{verify}: the anchor is intact on 1.18286.0 (var \code{ms}, \code{||hide()} present, no \code{blur()}), but the defect lives in Electron's Linux \code{isFocused()}, so unchanged bytes cannot distinguish ``still broken'' from ``Electron fixed it.'' One verdict firms without flipping: \code{org-plugins} goes from survivor candidate to a settled \bsurv{}, its byte-gap confirmed on pristine 1.18286.0 (\code{default:return null}, no linux case). A \bverify{} badge is an annotation of residual risk, \textbf{not} a code reversal: \code{frame-fix} and \code{wco-shim} are deleted and shipped, and only \code{quick-window} and \code{org-plugins} sit in \code{active\_patches}.}

\unithead{Ground-truth corrections}

\reppar{The verification pass ran against the reassessment as much as against the audit, and corrected one of its claims. The reassessment framed the launcher's dropped password-store \emph{auto-detection} as a deleted regression and a hard blocker; that framing is wrong and is deliberately kept out of this report. \code{--password-store} still passes when \code{CLAUDE\_PASSWORD\_STORE} is set (a documented escape hatch citing \#593), the doctor still reports the backend, and what actually changed is that the official \code{os\_crypt} autodetect now owns the default, a deliberate documented rework. The one genuinely open part is \textbf{LD-1}: whether that autodetect persists cookies on a live KDE/kwallet6 session, which static analysis cannot settle. The pass also surfaced two byte facts the audit had not called out. The official tree is bare co-located with \textbf{no \code{node\_modules/electron/dist} directory}, so the three artifact-test scripts that still asserted that on-disk path (\textbf{SB-1}) failed against the rebase packages --- now repointed to the bare layout (the companion \code{.bats} references were only synthetic strings fed to path-agnostic matchers and never failed); see the settlement below. And the payload compression changed across the train, \code{data.tar.zst} on 1.17377.2 to \code{data.tar.xz} on 1.18286.0, both already handled by \code{\_extract\_deb\_member}.}

\unithead{Live-hardware settlement (Rev B, July 3--4, 2026)}

\reppar{The open checks were then run on the local libvirt test fleet and a real KDE Plasma/kwallet6 host against the rebased 1.18286.0 build, and the two deletions carrying the most residual risk are now confirmed. \textbf{FF-1 is resolved}: on a focus-follows-mouse compositor (niri) the deleted \code{webContents.focus()} sloppy-WM guard produces no anomalous focus-steal (\#416); the \#605 sleep inhibitor both registers \emph{and} releases at the D-Bus layer on KDE (a session-bus \code{Inhibit}/\code{UnInhibit} capture paired every cookie), so the never-released-inhibitor bug does not reproduce --- though keep-awake is a silent no-op on bare wlroots/i3, which expose no \code{SessionManager} or \code{PowerManagement} inhibit service (a separate functional gap, upstream candidate); \code{openAtLogin} (\#128) survives a reboot; and quit-without-tray (\#321/\#623) is handled natively by the official build's \textbf{Settings \textrightarrow{} General \textrightarrow{} System Tray} toggle (off = quit-on-close), so the deleted \code{CLAUDE\_QUIT\_ON\_CLOSE} hatch is superseded, not lost. \textbf{WCO-1 is resolved}: the in-app claude.ai topbar renders on both KDE and niri, so the server-delivered \code{isWindows()} UA gate does not hide it on Linux and \code{wco-shim} needs no UA-override survivor (sway draws a second, server-side titlebar over it --- a decoration-suppression bug, not a topbar-absence one). \textbf{The \#400 config-merge check is closed as a deliberate no-op}: the loss reproduces only when the config file is edited \emph{while the app runs}; the normal edit-then-restart flow is safe, and extraction of the 1.18286.0 bundle showed \code{setMcpServers} now deletes entries programmatically (\code{delete i[a]}), so the retired \code{Object.assign} merge would resurrect a legitimately-deleted server --- the patch correctly stays out of \code{active\_patches}. \textbf{SB-1 is fixed} (repointed as above). The verification also surfaced findings outside the reassessment's frame, the load-bearing one being \textbf{LOG-1}: the launcher logged the full argv, so a relaunch through the login redirect wrote the \code{claude://login/\ldots?code=} OAuth authorization code to \code{launcher.log} in plaintext --- now redacted at the \code{log\_message} chokepoint.}

\unithead{What still needs hardware}

\reppar{Two checks narrow rather than close. \textbf{QW-1} confirmed the happy-path Quick Entry submit-then-raise flow on both the patched (KDE) and unpatched (niri) paths, but the KDE stale-focus raise edge --- does the popup reliably reappear when the main window is hidden or visible-but-unfocused --- still decides whether \code{quick-window} stays. \textbf{LD-1} passes on a KDE session with a pre-existing wallet (the official \code{os\_crypt} autodetect seals cookies, auto-probe dropped) but leaves the fresh-no-wallet freeze edge untested; keyring-less compositors persist via the \code{basic} backend, storing the token unencrypted at rest behind an advisory prompt. Two subsystem checks stay genuinely open and off the 3.0.0 critical path: live arm64 rootfs availability, and a Cowork socket-protocol capture on a KVM host (feeding the 3.1 \code{cowork-bwrapd} scoping). Separately, the no-hardware follow-ons ride their own pull requests: \textbf{ACQ-1} (the Nix derivation is still a hard \code{throw}, and Nix is the single largest install channel in the catalog), \textbf{LD-2} (note \code{CLAUDE\_QUIT\_ON\_CLOSE} as a no-op in the doctor's legacy-env list, now that the tray toggle supersedes it), and \textbf{AU-1}/\textbf{MB-1} (build tripwires on \code{apt\_channel\_pending} and \code{menuBarEnabled:!0} so a future upstream flip is caught at build time). The settled verdicts are the \href{https://github.com/aaddrick/claude-desktop-debian/blob/main/docs/learnings/official-deb-rebase-verification.md}{patch-necessity matrix}; the reassessment's full reasoning is the \href{https://github.com/aaddrick/claude-desktop-debian/blob/main/docs/reports/CDL-ANT-0009_patch-suite-history/verdict-reassessment-accessibility.md}{accessibility reassessment}.}
